Conservation decision-making under long transients

### 2019-08-12 slides/solarized.Rmd
solarized theme

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev.args=list(bg="transparent"), 
                      echo = FALSE, message=FALSE, warning=FALSE,
                      fig.width=11, fig.height=6.5, cache = TRUE)

library(tidyverse)
library(ggthemes)

theme_set(
 theme_bw(base_size=26) + 
  theme(line = element_line(colour = "white", size = 1.5),
        rect = element_rect(fill = "transparent", colour = "white"),
        text = element_text(colour = "white"),
        axis.text = element_text(colour = "white"),
        axis.ticks = element_line(colour = "white", size = 1.5),
        panel.border = element_rect(colour = "white", size = 1.5),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_blank(), 
        legend.background = element_blank(), 
        legend.key = element_rect(fill = "transparent", colour = "transparent"), 
        strip.background = element_blank(),
        strip.text = element_text(colour = "white"),
        panel.grid = element_blank()))

source("../R/ghost.R")
```

@@

@@

---
background-image: linear-gradient(to right, rgba(150, 150, 150, .1), rgba(150, 150, 150, .4)), url(img/city-mit-license.jpg)
background-size: 140%
class: center, top, inverse

@@

---
name: polarbear
background-image: linear-gradient(to right, rgba(150, 150, 150, .1), rgba(150, 150, 150, .4)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, top, inverse

---
template: polarbear

@@

---
layout: true
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/whale-shark.jpg)
background-position: center
background-size: 120%
class: center, top, inverse

@@

![](img/ghost-invert.png)

@@

# Ghost Attractor

@@

```{r}
data %>% filter(reps == lows[[6]]) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(lwd=2, col="white") + 
  ylim(0,1.6)
```

@@

```{r arima_data}
arima_forecast %>% filter(t < 2000) %>%
  ggplot(aes(t)) + 
  geom_line(aes(y=x), col="white") +
  geom_line(aes(y=pred), col="grey", lwd = 1) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se), fill="white", alpha = 0.5) + 
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

```{r arima_forecast1}
arima_forecast %>% ggplot(aes(t)) + 
  geom_line(aes(y=pred), col="grey", lwd = 1) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se), fill="white", alpha = 0.5) + 
  xlim(0,3000) + ylim(0, 1.6) + ylab("x")
```

@@

```{r arima_forecast}
arima_forecast %>% ggplot(aes(t)) + 
  geom_line(aes(y=x), col="white") +
  geom_line(aes(y=pred), col="grey", lwd = 1) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se), fill="white", alpha = 0.5)  + 
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

---
layout: true
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .9)), url(img/royalsociety/snake.jpg)
background-size: 120%
background-position: center
class: left, top, inverse

@@

```{r theory_U}
theory(p) %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd = 2, col="white")

@@

```{r det}
det <- det_sim(may,p) %>% mutate(reps=1) %>% filter(t < 3000)
det  %>%
  ggplot(aes(t, x)) + 
  geom_line(lwd = 2, col="white")
```
---
layout: true
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .95)), url(img/royalsociety/gentoo-penguin.jpg)
background-size: 150%
class: center, top, inverse
background-position: left

@@

@@

```{r stoch1}
data %>% filter(reps == lows[[6]]) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(lwd=2, col="white") + 
  ylim(0,1.6)
```

@@

data %>% filter(reps == highs[[6]]) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(lwd=2, col="white") + 
  ylim(0,1.6)

@@

```{r stoch_all}
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white")
```

@@

```{r ensemble}
mean <- data %>% group_by(t) %>% summarise(x = mean(x)) %>% mutate(reps = 1)
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white") + 
  geom_line(data = mean, lwd=2, color = "white")
```

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .7), rgba(50, 50, 50, .95)), url(img/royalsociety/gentoo-penguin-wtf.jpg)

```{r wtf}
data %>% 
  ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02, color = "white") + 
  geom_line(data = mean, lwd = 2, color = "white") +
  geom_line(data = det, lwd = 2, color = rgb(.6, .6, 1)) 
```

@@

@@

```{r fpt}
passage %>% ggplot(aes(tau)) + 
  geom_histogram(binwidth = 200, fill=rgb(.9, .9, .9, .7)) + 
  geom_vline(aes(xintercept=det_tau), col="blue", lwd=2)
```

---
layout: false
background-image: linear-gradient(to right, rgba(50, 50, 50, .7), rgba(50, 50, 50, .95)), url(img/royalsociety/green-crested-lizard.jpg)
background-size: 110%
class: left, top, inverse
background-position: center

@@

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .8), rgba(50, 50, 50, .9)), url(img/royalsociety/reef-shark.jpg)
background-size: 120%
class: top, inverse
background-position: center
layout: true

@@

@@

```{r disc_sims}
discrete_sims <- read_csv("../data/discrete-sims.csv.xz")
discrete_sims %>% filter(series %in% c("1", "2")) %>%
  ggplot(aes(time,state)) + geom_line(col="white", lwd=1.5) + facet_wrap(~ series)
```

@@

@@

```{r posteriors, fig.width=13, fig.height=6}
label <- c("no switch", "switch", "no ghost")
pal <- viridis::viridis_pal()(3)
df <- c("../data/no_switches.csv.xz", "../data/switches.csv.xz", "../data/weak_ghost.csv.xz") %>%
  map_dfr(read_csv, .id = "timeseries") %>% 
  mutate(timeseries = label[as.integer(timeseries)])

inset <- discrete_sims %>% mutate(series = label[series]) %>% 
  ggplot(aes(time,state, col=series)) + 
  geom_line(lwd=1.5, show.legend = FALSE) + 
  ggplot2::scale_color_viridis_d()

p1 <- df %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=2) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=2) + 
  ggplot2::scale_fill_viridis_d()

vp <- grid::viewport(width = .5, height = 0.55, x = 0.35, y = 0.7)

@@

---
layout: true
background-image: linear-gradient(to right, rgba(30, 30, 30, .8), rgba(30, 30, 30, .9)), url(img/royalsociety/forest-fire.jpg)
background-size: 120%
class: center, top, inverse

@@

```{r}
library(MDPtoolbox)
library(sarsop) # remotes::install_github("boettiger-lab/sarsop")
library(tidyverse) # for plotting
library(mdplearning)

@@

possible_a <- seq(.25, .34, by = 0.005)
true_a <- 0.27   ## reality has just a transient.  use  0.277 for stronger ghost
believe_a <- 0.31 ## believe there's a strong attractor
true_i <- which.min(abs(possible_a - true_a))
```

**True model: weak ghost attractor**

```{r true_a}
f <- may(true_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col="white") +
  geom_point(col="white") + 
  geom_hline(aes(yintercept = 0),col="white")
```

@@

```{r a29}
f <- may(0.29)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col="white") +
  geom_point(col="white") + 
  geom_hline(aes(yintercept = 0),col="white")
```

@@

```{r a31}
f <- may(believe_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col="white") +
  geom_point(col="white") + 
  geom_hline(aes(yintercept = 0), col="white")
```

@@

```{r}
m_true <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(true_a), sigma_g, sigma_m, noise = "lognormal")

@@

**no Intervention**

@@

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) +
  geom_point(col="white")+
  geom_line(aes(time, action), col="grey", lwd=2) + ylim(0,1.4)
```

@@

```{r priors}
prior <- dnorm(possible_a, 0.29, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", fill="white") +
  geom_vline(aes(xintercept = true_a), col="white", lwd=1, lty=2) 
```

@@

```{r priors_wide}
prior <- dnorm(possible_a, 0.29, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", fill="white") +
  geom_vline(aes(xintercept = true_a), col="white", lwd=1, lty=2) 
```

@@

### Less learning, more caution

---
layout: true
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/pensive-polarbear.jpg)
background-size: 120%
class: center, top, inverse

---
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/eagle-ray.jpg)
background-size: 120%
class: center, top, inverse

---
class: center, middle
background-image: linear-gradient(to right, rgba(50, 50, 50, .6), rgba(50, 50, 50, .9)), url(img/royalsociety/bohemian-waxwing.jpg)
background-size: 120%

@@

![](img/royalsociety.jpg) Photo Competition 
]
.pull-right[

>>>>

---
title: "Conservation decision-making under long transients"
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["default","solarized-light.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: "solarized-light"
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"

@@

# Conservation decision-making under long transients

## Carl Boettiger

## UC Berkeley
---
layout: true
background-color: #fdf6e3
class: center, top

```{r setup, include=FALSE}
knitr::opts_chunk$set(#dev.args=list(bg="transparent"), 
                      echo = FALSE, 
                      message=FALSE, 
                      warning=FALSE,
                      fig.width=11, 
                      fig.height=6.5, 
                      cache = TRUE)

library(tidyverse)
library(ggthemes)
theme_set(theme_solarized(base_size=26))
scale_colour_discrete <- function(...) scale_colour_solarized()
scale_fill_discrete <- function(...) scale_fill_solarized()
pal <- solarized_pal()(3)

txtcolor <- "#586e75"
```

```{r}
source("../R/ghost.R")
```

@@

![](img/royalsociety/penguin-swim.jpg)

@@

<div style='float:left;width:50%'>
  <img src='img/city-mit-license.jpg' style='transform:translate(-20%, -10%);height:100%;object-fit:cover;overflow:hidden'>
</div>

@@

<div style='float:left;width:50%'>
  <img src='img/royalsociety/pensive-polarbear.jpg' style='transform:translate(-20%, -10%);height:900px;overflow:hidden'>
</div>

@@

@@

@@

@@

```{r}
data %>% filter(reps == lows[[6]]) %>%
  ggplot(aes(t, x, group=reps, col=as.character(reps))) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  ylim(0,1.6) + scale_color_solarized(accent="blue")
```

@@

```{r arima_data}
arima_forecast %>% filter(t < 2000) %>%
  ggplot(aes(t)) + 
  geom_line(aes(y=x),  lwd = 1, col=pal[1]) +
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

```{r arima_forecast1}
arima_forecast %>% ggplot(aes(t)) + 
  geom_line(aes(y=pred),  lwd = 1.5, col=pal[1]) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se),  alpha = 0.5, fill=pal[1]) + 
  xlim(0,3000) + ylim(0, 1.6) + ylab("x")
```

@@

```{r arima_forecast}
arima_forecast %>% ggplot(aes(t)) + 
  geom_line(aes(y=x), col=pal[2], lwd=1) +
  geom_line(aes(y=pred),  lwd = 1.5, col=pal[1]) + 
  geom_ribbon(aes(y=pred, ymin=pred-se, ymax=pred+se),  alpha = 0.5, fill=pal[1])  + 
  xlim(0,3000) + ylim(0, 1.6)+ ylab("x")

@@

@@

```{r theory_U}
theory(p) %>%
  ggplot(aes(x, potential)) + 
  geom_line(lwd = 2,col=txtcolor)

@@

```{r det}
det <- det_sim(may,p) %>% mutate(reps=1) %>% filter(t < 3000)
det  %>%
  ggplot(aes(t, x)) + 
  geom_line(lwd = 2,col=txtcolor)
```

@@

```{r stoch1}
data %>% filter(reps == lows[[6]]) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(lwd=2,col=txtcolor) + 
  ylim(0,1.6)
```

@@

@@

data %>% filter(reps == highs[[6]]) %>%
  ggplot(aes(t, x, group=reps)) + 
  geom_line(lwd=2,,col=txtcolor) + 
  ylim(0,1.6)

@@

```{r stoch_all}
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02,col=txtcolor)
```

@@

```{r ensemble}
mean <- data %>% group_by(t) %>% summarise(x = mean(x)) %>% mutate(reps = 1)
data %>% ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02,col=txtcolor) + 
  geom_line(data = mean, lwd=2,col=pal[1])
```

@@

```{r wtf}
data %>% 
  ggplot(aes(t,x, group=reps)) + 
  geom_line(alpha=0.02,col=txtcolor) + 
  geom_line(data = mean, lwd = 2, col=pal[1]) +
  geom_line(data = det, lwd = 2, col=pal[2]) 
```

@@

```{r fpt}
passage %>% ggplot(aes(tau)) + 
  geom_histogram(binwidth = 200, col=txtcolor) + 
  geom_vline(aes(xintercept=det_tau), col=pal[1], lwd=2)
```

@@

@@

![](img/royalsociety/reef-shark.jpg)

@@

```{r disc_sims}
discrete_sims <- read_csv("../data/discrete-sims.csv.xz")
discrete_sims %>% filter(series %in% c("1", "2")) %>%
  ggplot(aes(time,state)) + 
  geom_line(lwd=1.5, col=pal[1]) + 
  facet_wrap(~ series)
```

@@

# Model Inference

@@

# Model Inference

```{r posteriors, fig.width=13, fig.height=6}
label <- c("no switch", "switch", "no ghost")
df <- c("../data/no_switches.csv.xz", "../data/switches.csv.xz", "../data/weak_ghost.csv.xz") %>%
  map_dfr(read_csv, .id = "timeseries") %>% 
  mutate(timeseries = label[as.integer(timeseries)])

inset <- discrete_sims %>% mutate(series = label[series]) %>% 
  ggplot(aes(time,state, col=series)) + 
  geom_line(lwd=1.5, show.legend = FALSE)

p1 <- df %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9, show.legend = FALSE) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=2) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=2)

vp <- grid::viewport(width = .4, height = 0.45, x = 0.35, y = 0.7)

@@

@@

```{r results="hide"}
library(MDPtoolbox)
library(sarsop) # remotes::install_github("boettiger-lab/sarsop")
library(tidyverse) # for plotting
library(mdplearning)

@@

possible_a <- seq(.25, .34, by = 0.005)
true_a <- 0.27   ## reality has just a transient.  use  0.277 for stronger ghost
believe_a <- 0.31 ## believe there's a strong attractor
true_i <- which.min(abs(possible_a - true_a))

## True model: weak ghost attractor

```{r true_a}
f <- may(true_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col=pal[1], lwd=2) +
  geom_hline(aes(yintercept = 0), lwd=2, lty=2) + 
  ylim(-0.06, 0.06)
```

@@

```{r a29}
f <- may(0.29)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col=pal[1], lwd=2) +
  geom_hline(aes(yintercept = 0), lwd=2, lty=2) + 
  ylim(c(-0.04, 0.04))
```

@@

```{r a31}
f <- may(believe_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + 
  geom_line(col=pal[1], lwd=2) +
  geom_hline(aes(yintercept = 0), lwd=2, lty=2) + 
  ylim(-0.03, 0.03)
```

@@

```{r results="hide"}
m_true <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(true_a), sigma_g, sigma_m, noise = "lognormal")

@@

## Simulation under *no intervention*

@@

df %>% mutate(state = states[state], action = actions[action]) %>% 
  select(state, action, time) %>% 
  gather(series, state, -time) %>%
  ggplot(aes(time, state, col=series)) +
  geom_line(lwd=2) + 
  ylim(0,1.4)
```

@@

```{r priors}
prior <- dnorm(possible_a, 0.29, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", col=txtcolor) +
  geom_vline(aes(xintercept = true_a), lwd=1, lty=2, col=pal[1]) 
```

@@

```{r priors_wide}
prior <- dnorm(possible_a, 0.3, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", col=txtcolor) +
  geom_vline(aes(xintercept = true_a), lwd=1, lty=2, col=pal[1]) 
```

@@

## Less learning, more caution

```{r}
```

---
class: center, top
background-image: url(img/royalsociety/bohemian-waxwing.jpg)
background-size: 140%

# Thanks!

@@

![](img/royalsociety.jpg)

**Publishing Photo Competition**
